---
title: "Predictive Machine Learning Project"
author: "Subhash"
date: "Sunday, October 25, 2015"
output: html_document
---

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(rattle)
```

```{r}
training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))

Test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))
```

Delete columns where entire column is NA
```{r}
training<-training[,colSums(is.na(training)) == 0]
Test <-Test[,colSums(is.na(Test)) == 0]
```

Remove first 7 columns since they dont contain any useful info that require prediction

```{r}
training   <-training[,-c(1:7)]
Test <-Test[,-c(1:7)]
```

Create partition within Training for cross validation
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
sub_training <- training[inTrain,]
sub_testing <- training[-inTrain,]
```

##Prediction using Rpart
```{r}
#model training
model1<- rpart(classe~., data=sub_training, method="class")
rpart.plot(model1)

#prediction using the model
prediction1 <- predict(model1, sub_testing, type="class")

#Confusion matrix to test the prediction
confusionMatrix(prediction1, sub_testing$classe)

```

```{r, echo=FALSE, results='hold'}
gc()
```
We see tthat the accuracy is low. Now lets try with Random Forest model

##Prediction using Random Forest
```{r}
#model training
model2<- randomForest(classe~., data=sub_training, method="class")

#prediction using the model
prediction2 <- predict(model2, sub_testing, type="class")

#Confusion matrix to test the prediction
confusionMatrix(prediction2, sub_testing$classe)
```
As expected the Random Forest model has better accuracy as compared to rPart model. So, we chose Random Forest model to apply on the 20 Test samples

##Apply Random Forest model on the 20 Test records
```{r}
predict_actual <- predict(model2, Test, type="class")

#create an output table
output <- data.frame(ProblemID = numeric(0), classe=numeric(0))
p<-as.character(predict_actual)
for (i in 1:length(p)){
        output[i,1]<-i
        output[i,2]<-p[i]     
        
}

output

```
##Write files for submission
```{r}
# Write files for submission
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predict_actual)
